---
title: "basics_R"
author: "Binod"
date: "6/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Your Turn 01

## Assign number in num1  and print out

```{r}
num1 <- 

## print num1
print()  
```

## Assign alphabet in alpha1 and print out

```{r}
alpha1 <-
  
print()
```

## Naming Convention

```{r}
mydata <- 234   ## correct
 
# 1_my_data <- 345  ## in correct

my_data_1 <- 456 ## correct
```

### Case sensetive

```{r}
mydata <- 234

MyData <- 567

print(mydata)

print(MyData)
```

## Data Types

```{r}
a <- 100
alp <- "alphabet"

```

```{r}
class(a)

class(alp)

class(mydata)
```


## Find Data Type Present in Type 1

```{r}
# class of num1

class()
```

```{r}
# class of alpha1s
class(alpha1)
```

## Data Structures

- Vectors
- Lists
- Matrices
- Arrays
- Factors
- Data Frames


## Vectors

- Vector is collection of same type of values
- created using c() function

```{r}
## create a new vector

vec1 <- c(1, 2, 3)

vec1

class(vec1)
```

```{r}
vec2 <- c("ram", "shyam", 1)

vec2

class(vec2)
```


# Your Turn 3

- create a new vector
- find the data type of vector

```{r}

```

## Opearators

- Arithmetic Operators
- Relational Operators
- Logical Operators


### Arithmetic Operators


#### Addition

```{r}
# Addition
a <- c(5, 6, 7)
b <- c(2, 1, 3)

print()
```

### Subtraction

```{r}
#subtraction
a <- c(9,8,7)
b <- c(2,4,9)

print()
```

#### Multiply

```{r}
# Multiply

a <- c(1, 2, 3)
b <- c(2, 3, 4)

print()
```

#### Division

```{r}
# Division
a <- c(22,13,24)
b <- c(2,3,4)

print(a / b)
```

### Exponent

```{r}
4 ^ 2

99 ^ 3
```

### Modulus

```{r}
7 %% 2

99 %% 8
```

## Relational Operators

### Greater than ( > )

```{r}
# greater than
a = c(5,2,3,4)
b = c(10,1,2,9)

print()
```

### Less Than ( < )

```{r}
#less than <
a <- c(9,10,11,21)
b <- c(11,2,3,55)

print(a < b)
```

### Equals to ( == )

```{r}
#equals to ==

a <- c(2,3,4)
b <- c(2,3,5)
print()
```

### NotEqual ( != )

```{r}
# notequal to
a <- c(2,3,4,10)
b <- c(1,3,4,9)

print()
```

### Less than equal to (<=)

```{r}
#less than equal to
a <- c(2,3,4,9)
b <- c(2,5,7,8)

print()
```

### Greater than equal to(>=)

```{r}
# greater than equal to
a <- c(2,3,4,10)
b <- c(5,7,4,9)

print( )
```


## Logical Operators

### AND operator 

```{r}
# AND operator
a <- c(3,9,5)
b <- c(5,6,7)
print(a>5 & b>5)
```

### OR operator

```{r}
# OR operator
a <- c(1,5,3)
b <- c(4,5,6)

print()
```


### Logical NOT ( ! )

```{r}
# logical not
a <- c(3,4,5)

print()
```


# DataFrame

```{r}
mtcars
```


## Lets Create Dataframe

```{r}
data <- data.frame(id = c(1:5), 
                   name = c("Richa","Deepak","Manisha","Ryan","Ram"),
                   age = c(23,25,45,30,57),
                   gender = as.factor(c("female","male","female","male","male")),
                   stringsAsFactors = FALSE)

data
```
## Get the Structure of the Data Frame

```{r}
# Get the structure of the data frame.
str(data)
```
## Summary of Data in Data Frame

```{r}
# Print the summary
summary(data)
```
## Extract Data From DataFrame

*head*

```{r}
head()
```

*tail*

```{r}
tail()
```

## Extract Data

- Note: Index count starts from 1 (not 0)


> Extract first two rows and first two columns

```{r}
data[]
```
> Extract 2nd row and all columns

```{r}
data[,]
```

> Extract 3rd row with 2nd and 4th column

```{r}
data[, c()]
```

> Extract 2nd and 4th row with 2nd, 3rd and 4th column

```{r}
data[c(), c()]
```

# Your Turn 4
- Extract name and age column
- Extract 2nd row
- Extract 3rd and 4th row and name, gender,age column

```{r}

```

## Add Column

```{r}
#Add the 'address' column
data$address <- c("butwal","kathmandu","brigunj","illam","karnali")

data
```

## Your Turn 5

- Add "blood_group" column to the dataframe

```{r}
data$blood_group <- c()
data
```

## Add Row

We can add rows to dataframe using rbind() function.


```{r}
# Create second data frame
newdata <- data.frame(id = c(6:8),
                name = as.character(c("Deepika","Manish","Brion")),
                age = c(33,15,40),
                gender = as.factor(c("female","male","male")),
                address = c("patan","dharan","butwal"),
                blood_group = c("A+","B+","AB-"), 
                stringsAsFactors = FALSE)

newdata
```

```{r}
final_data <- rbind(data,newdata)
final_data
```

## Merge two dataframe

```{r}
prof_data <- data.frame(
   name = as.character(c("Deepika", "Manish", "Brion", "Richa", "Deepak", "Manisha", "Ryan", "Ram")), 
   phone = c(9129345550, 9129365570, 9129346789, 9129343457, 9129342344, 9129343451,
9129344670, 9129345650), 
   profession = as.character(c("Doctor", "Nurse", "Driver", "Scientist", "Computer
Technician", "Teacher", "Student", "Lawyer")),
stringsAsFactors = FALSE)
```

```{r}
merge(prof_data, final_data, by = "name")
```

## Rename Column

- select column with its name and replace with new name
- name() function gives the name of the dataframe column

```{r}
names(final_data)[names(final_data)=="name"] <- "first_name"
final_data
```


## Function

```{r}
# function for adding number
add_me <- function(num1,num2){
sum <- num1 + num2
return(sum)
}

add_me(99,123)
```


Built In function
R has a large number of in-built functions

```{r}
# Create a sequence of numbers from 32 to 44.
print(seq(32,44))
```


```{r}
# Find mean of numbers from 25 to 82.
print(mean(25:82))
```

```{r}
# Find sum of numbers frm 41 to 68.
print(sum(41:68))
```

# Install Package

```{r}
install.package() # is used to install package.
install.packages("package name")
library()  # is used to load the package.
library("package name")
```

## Asking for Help

```{r}

# asking help about the mean() function
?mean

# asking help about the mtcars datasets
?mtcars

# asking help about the dplyr package
?dplyr
```

### Reference Book

https://rstudio-education.github.io/hopr

